The Heavy Photon Search Test Detector

### 2013-10-24 hps-testrun-nim.tex
Small update from Pelle

% HPS stuff
%\usepackage{multicol}
\newcommand{\Aprime}{A$^\prime$}
\newcommand{\ee}{e$^+$e$^-$}

@@

There are two necessary parameters that determine the properties of this ``heavy photon," designated the 
\Aprime.  The first is the mixing of this state with the Standard Model photon, $\epsilon$, which results in 
a coupling of the \Aprime to Standard Model fermions with a strength of $\epsilon e$.  In simple models 
with kinetic mixing, $\epsilon$ is favored to be in the range of $10^{-2} -10^{-5}$.  The second is the mass 
of the \Aprime.  Under the assumption that this mass is generated via a Higgs-like symmetry breaking, 
the mass would be roughly of the order $\sqrt{\epsilon} m_W \sim$~MeV-GeV.  These ranges are further 
motivated by explanations of other dark matter anomalies involving non-minimal dark sectors~
\cite{TuckerSmith:2001hy} and the fact that no similar excess in proton-antiproton pairs is observed, which 
would be simply explained if $ m_{\textrm{A}^\prime} < 2m_p$~\cite{Adriani:2010rc}.

@@

It is this last challenge that has the most interesting and obvious impacts on the design of the SVT. To 
achieve good vertexing performance, the first layer must be placed no more than about 10 cm downstream 
of the target. At that distance, it is found that the active region of a sensor can be placed as close to 1.5 mm 
from the center of the beam, defining a 15 mrad ``dead zone" in the detector around the plane of the 
primary beam. At the edge of this ``dead zone," the radiation dose approaches $10^{15}$ electrons/cm
$^2$/month, or roughly $3 \times 10^{13}$ 1 MeV neutron equivalent/cm$^2$/month, as shown in Fig.~
\ref{fig:radiation} and~\ref{fig:rates}~\cite{Rashevskaya:2002nd}. 
%===================
\begin{figure}[htp]
\includegraphics[width=7cm]{figures/radiation}
\label{fig:radiation}
\caption{Electron fluences: electrons bend towards $+x$.}
\end{figure}
\begin{figure}[htp]
\includegraphics[width=7cm]{figures/rates}
\label{fig:rates}
\caption{The environment near the dead zone 10 cm from the target.}
\end{figure}
%\begin{figure}[htp]
%\subfigure[\label{fig:rates}The occupancy in an 8 ns window in 60 micron strips.]
%{\includegraphics[width=8cm]{figures/rates}}
%\subfigure[\label{fig:radiation}Electron fluences: electrons bend towards $+x$.]
%{\includegraphics[width=7cm]{figures/radiation}}
%\subfigure[\label{fig:rates}The occupancy in an 8 ns window in 60 micron strips.]
%{\includegraphics[width=8cm]{figures/rates}}
%\caption{The environment near the dead zone 10 cm from the target.}
%\end{figure}
%===================
This requires the sensors to be actively cooled. Meanwhile, very low-energy delta rays from beam gas interactions multiply the density of background hits, so the SVT must operate inside the beam vacuum.  Finally, in order protect the sensors, the detector must be movable so that it can be retracted during periods of uncertain beam conditions.

%To summarize the SVT must:
%\begin{itemize}
%\item have a uniform acceptance in the forward region; up to $\theta_y=XXXXmrad$ and as close as 
%possible to the beam axis to catch boosted decay products,
%\item >95\% reconstruction efficiency for electron and positrons  with energy down to $XXXX$~MeV,
%\item momentum resolution of XXXX\% to separate a promptly decaying \Aprime signal from the irreducible background,
%\item vertex resolution $<XXXX$~mm along the beam line direction to search for displaced \Aprime decays,
%\item event timing resolution down to 2~ns, 
%\item sustain a occupancy of up to 1\%, trigger rates of $\approx50$~kHz and data rates up to 100~MB/s,
%\item survive and efficiently operate after $1\times10^{14} neutron eq. fluence,
%
%dosedetector components that can survive and efficiently operate in a high radiation environment with 
%local doses exceeding $XXXX$~MRad
%\end{itemize}

@@

A realistic simulation of the detector using the java-based org.lcsim package and including full backgrounds has been used to optimize the layout for acceptance, tracking efficiency, mass resolution and prompt vertex rejection~\cite{Graf:2011zzc}. The layout of the SVT is summarized in Table~\ref{tab:trk} and shown in Figure~\ref{fig:tracker_model}.

The layout of the SVT is shown in Fig.~\ref{fig:tracker_model} and summarized in Table \ref{tab:trk}. The 
vacuum chamber is too small vertically to accommodate 90-degree stereo layers to optimize the vertex 
resolution. Instead, 100~mrad stereo is used in the first three layers 
to provide higher-resolution 3-D space points for vertexing. The 50~mrad stereo of the last two layers 
breaks the tracking degeneracy of having five identical layers and minimizes fakes from ghost hits, 
improving pattern recognition while still providing sufficient pointing resolution into the third layer
 for robust hit 
association in the denser environment there. Altogether, this layout comprises 20 sensors and hybrids and 
100 APV25 chips for a total of 12780 readout channels. 
%=======================
\begin{center}
\begin{table}[ht]
{\small
\begin{tabular}{lccccc}   
\hline \hline 
    Layer & 1 & 2 & 3 & 4 & 5 \\      
\hline
    $z$ from target (cm)  & 10 & 20 & 30 & 50 & 70  \\ 
    Stereo angle (mrad)  & 100 & 100 & 100 & 50 & 50 \\ 
    Bend res. ($\mu$m)  & $\approx$60 & $\approx$60 & $\approx$60 & $\approx$120 & $\approx$120  \\ 
    Non-bend res. ($\mu$m)  & $\approx$6 & $\approx$6 & $\approx$6 & $\approx$6 & $\approx$6  \\ 
    \# of sensors  & 4 & 4 & 4 & 4 & 4  \\ 
    Dead zone (mm) & $\pm1.5$  & $\pm3.0$  & $\pm4.5$  & $\pm7.5$  & $\pm10.5$  \\ 
    Power cons. (W) & 6.9 & 6.9 & 6.9 & 6.9 & 6.9 \\
\hline \hline
\end{tabular}
\caption{Layout of the HPS Test SVT.  Note the elimination of one layer and 90-degree stereo, and the use of only small, single-sensor-sized double-sided modules both above and below he dead zone. }
}
\label{tab:trk} 
\end{table}
\end{center}
%=======================
\begin{center}
\begin{figure}[htp]
    \includegraphics[width=7cm]{figures/HPS_nochamber}
\caption{A rendering of the Test SVT showing the modules on their support plates held by the hinged C-support on the left and the motion levers on the right. The sensors are shown in red and the hybrids in green. The beam enters from the right through a vacuum box with flanges for services. }
\label{fig:tracker_model}
\vspace*{-5mm}
\end{figure}
\end{center}
%=======================
The Test SVT is built in two separate halves that are mirror reflections of one another about the plane of the 
nominal electron beam.  Each half consists of five, double-sided modules mounted on a support plate that 
provides services to the modules and allows them to be moved as a group relative to the dead zone. The 
two halves of the tracker are connected to hinges mounted on a C-shaped support just beyond layer 5 that 
defines the nominal spacing between the upper and lower halves of the tracker.  A shaft attached to each 
support plate in front of layer 1 extends upstream and connects to a linear shift that transfers motion into the 
vacuum box through bellows to open and close the tracker around the dead zone. The C-support is 
mounted to an aluminum baseplate that defines the position of the SVT with respect to the vacuum 
chamber. Figure~\ref{fig:tracker_halves} shows a photograph of both completed detector halves prior to 
final assembly. 
%=======================
\begin{figure}[htp]
\begin{center}
    \includegraphics[width=7cm]{figures/2012-101-PHOTON-DETECTOR-001}
\caption{\small{Both halves of the HPS Test SVT after final assembly at SLAC.  The cooling manifolds and 
integrated cable runs are clearly seen.} }
\label{fig:tracker_halves}
\end{center}
\vspace*{-5mm}
\end{figure}
%=======================

@@

\subsubsection{Silicon Sensors}
The sensors for the SVT are selected according to a number of requirements.  First, the sensors must be 
radiation tolerant to approximately $1.5\times10^{14}$ 1 MeV neq/cm$^2$ for a six month run.  This 
corresponds to about 4 MHz/mm$^2$, so the sensor and readout technology must be capable of handling 
very high local occupancies. Since sensitivity is limited by multiple scattering, a material budget of less 
than 1\% $X_0$/layer is imperative and far less is desirable.  The best vertex resolution with a small beam 
spot ($\approx 10 \mu$m) requires single-hit resolution better than 20 $\mu$m in both measurement 
coordinates.  Finally, the sensor technology must be mature and readily available at low cost.  Surprisingly, 
we find that silicon microstrips are the best match to these requirements, and a supply of appropriate 
microstrip sensors purchased from the Hamamatsu Photonics Corporation for the Run 2b upgrade of D\O~
\cite{Denisov:2001aa} are readily available. These are $p+$ on $n$, single sided, AC coupled, polysilicon-
biased sensors fabricated on $<100>$ silicon and have 30(60) micron sense(readout) pitch over their 4 cm 
$\times$ 10 cm surface. Although a maximum bias of 350 V is specified, the vast majority are operable to 
1000V and therefore remain fully depleted to a dose of approximately $1.5\times10^{14}$ 1 MeV neq/cm
$^2$.

\subsubsection{Front-end Electronics}

Because the regions of high occupancy are small spots in two dimensions, only a short length of any one 
strip sees significant occupancy, and the strips in that region act as long pixels. However, the rates are still 
very high and lowering the peak occupancy in the sensors to approximately 1\% for tracking requires a 
trigger window and tagging of hit times to roughly 8 ns, as shown in Figure~\ref{fig:rates}.  The FADC 
readout for the ECal and muon system are capable of this, but achieving achieving $2\sigma$ efficiency for 
silicon hits then requires 2 ns time resolution for the hits in the SVT. This performance can be achieved with 
the APV25 readout ASIC developed for the CMS experiment at CERN~\cite{French:2001xb}.  When 
operated in ``multi-peak mode", the APV25 captures successive samples of the output of the shaper in sets 
of three.  By fitting the known $CR$-$RC$ shaping curve to these samples, the initial time of the hit can be 
determined to a precision of 2 ns for S/N$>$25, an achievable figure with our sensors if read out 
individually~\cite{Friedl:2009zz}.  For the SVT, six-sample readout and the shortest possible shaping time 
(35 ns) will be used to best distinguish hits that overlap in time.
The APV25 chips are hosted on simple FR4 hybrids since these hybrids and their cooling system are 
outside the tracking volume.  Along with the sensors, these are assembled into large multi-sensor modules, 
one for each tracking view.

@@

The sensor modules for the SVT consist of a pair of identical half-modules, sandwiched back-to-back around an aluminum cooling block at one end and a similar PEEK spacer block at the other. Figure \ref{fig:tracker_module} shows a prototype module assembly.
%=======================
\begin{figure}[htp]
	\begin{center}
   	 \includegraphics[width=7cm]{figures/IMG_5200}
	\caption{\small{A prototype module assembly (foreground) with the 50 mrad (left) and 100 mrad (right) module assembly fixtures in the background.  A pair of cooling blocks and a spacer block can be seen on the fixtures.} }
	\label{fig:tracker_module}
	\end{center}
\vspace*{-5mm}
\end{figure}
%=======================
The cooling block provides the primary mechanical support for the module as well as cooling via copper 
tubes pressed into grooves in the plates. The spacer block defines the spacing between the sensors at the 
far end of the module, stiffens the module structure, and improves the stability of the sensor alignments.

>>>>

% HPS stuff
%\usepackage{multicol}
\newcommand{\Aprime}{A$^\prime$}
\newcommand{\ee}{e$^+$e$^-$}
\newcommand{\fluenceunit}{1~MeV~neutron~equivalent/cm\ensuremath{^2}}

@@

There are two necessary parameters that determine the properties of this ``heavy photon," designated the 
\Aprime.  The first is the mixing of this state with the Standard Model photon, $\epsilon$, which results in 
a coupling of the \Aprime to Standard Model fermions with a strength of $\epsilon e$.  In simple models 
with kinetic mixing, $\epsilon$ is favored to be in the range of $10^{-2} -10^{-5}$.  The second is the mass 
of the \Aprime.  Under the assumption that this mass is generated via a Higgs-like symmetry breaking, 
the mass would be roughly of the order $\sqrt{\epsilon} m_W \sim$~-GeV.  These ranges are further 
motivated by explanations of other dark matter anomalies involving non-minimal dark sectors~
\cite{TuckerSmith:2001hy} and the fact that no similar excess in proton-antiproton pairs is observed, which 
would be simply explained if $ m_{\textrm{A}^\prime} < 2m_p$~\cite{Adriani:2010rc}.

@@

It is this last challenge that has the most interesting and obvious impacts on the design of the SVT. To 
achieve good vertexing performance, the first layer must be placed no more than about 10 cm downstream 
of the target. At that distance, it is found that the active region of a sensor can be placed as close to 1.5 mm 
from the center of the beam, defining a 15 mrad ``dead zone" in the detector around the plane of the 
primary beam. At the edge of this ``dead zone," the radiation dose approaches $10^{15}$ electrons/cm
$^2$/month, or roughly $3 \times 10^{14}$~\fluenceunit/month, as shown in Fig.~
\ref{fig:radiation} 
%===================
\begin{figure}[]
\includegraphics[width=7cm]{figures/radiation}
\label{fig:radiation}
\caption{Simulation of the fluence 10~cm downstream of the target in the active region of the SVT ($|Y|>0.15~$cm) looking downstream; electrons bend towards $+x$.}
\end{figure}
%===================
and~\ref{fig:rates}~\cite{Rashevskaya:2002nd}.
%===================
\begin{figure}[]
\includegraphics[width=7cm]{figures/rates}
\label{fig:rates}
\caption{Simulation of the occupancy in the SVT 10~cm downstream of the target as a function of the 
vertical distance from the beam center.}
\end{figure}
%=================== 
This requires the sensors to be actively cooled. Meanwhile, very low-energy delta rays from beam gas interactions multiply the density of background hits, so the SVT must operate inside the beam vacuum.  Finally, in order protect the sensors, the detector must be movable so that it can be retracted during periods of uncertain beam conditions.  
%To summarize the SVT must:
%\begin{itemize}
%\item have a uniform acceptance in the forward region; up to $\theta_y=XXXXmrad$ and as close as 
%possible to the beam axis to catch boosted decay products,
%\item >95\% reconstruction efficiency for electron and positrons  with energy down to $XXXX$~MeV,
%\item momentum resolution of XXXX\% to separate a promptly decaying \Aprime signal from the irreducible background,
%\item vertex resolution $<XXXX$~mm along the beam line direction to search for displaced \Aprime decays,
%\item event timing resolution down to 2~ns, 
%\item sustain a occupancy of up to 1\%, trigger rates of $\approx50$~kHz and data rates up to 100~MB/s,
%\item survive and efficiently operate after $1\times10^{14} neutron eq. fluence,
%
%dosedetector components that can survive and efficiently operate in a high radiation environment with 
%local doses exceeding $XXXX$~MRad
%\end{itemize}

@@

A realistic {\sc Geant4} simulation of the detector based on the {\sc SLIC}~\ref{slic} simulation 
package~\cite{Graf:2011zzc}
including all backgrounds has been used to optimize the layout for acceptance, tracking efficiency, mass 
resolution and prompt vertex rejection. The layout of the SVT is summarized in Tab.~\ref{tab:trk} and rendered in Fig.~\ref{fig:tracker_model}.

The layout of the SVT is shown in Fig.~\ref{fig:tracker_model} and summarized in Table \ref{tab:trk}.
The vertex resolution with an asymmetric beam spot of $\approx 200\times40 \mu$m requires hit 
resolutions better than XXXX (XXXX)~$\mu$m in the vertical (horizontal) measurement coordinate. 
In addition, the vacuum chamber is too small vertically to accommodate 90-degree stereo layers to optimize 
the vertex resolution in the horizontal coordinate. Instead, 100~mrad stereo is used in the first three layers 
to provide higher-resolution 3-D space points for vertexing. The 50~mrad stereo of the last two layers 
breaks the tracking degeneracy of having five identical layers and minimizes fakes from ghost hits, 
improving pattern recognition while still providing sufficient pointing resolution into the third layer
 for robust hit 
association in the denser environment there. Altogether, this layout comprises 20 sensors and hybrids and 
100 APV25 chips for a total of 12780 readout channels. 
%=======================
\begin{center}
\begin{table}[ht]
{\small
\begin{tabular}{lccccc}   
\hline \hline 
    Layer & 1 & 2 & 3 & 4 & 5 \\      
\hline
    $z$ from nom. target (cm)  & 10 & 20 & 30 & 50 & 70  \\ 
    Stereo angle (mrad)  & 100 & 100 & 100 & 50 & 50 \\ 
    Bend res. ($\mu$m)  & $\approx$60 & $\approx$60 & $\approx$60 & $\approx$120 & $\approx$120  \\ 
    Non-bend res. ($\mu$m)  & $\approx$6 & $\approx$6 & $\approx$6 & $\approx$6 & $\approx$6  \\ 
    \# of sensors  & 4 & 4 & 4 & 4 & 4  \\ 
    Dead zone (mm) & $\pm1.5$  & $\pm3.0$  & $\pm4.5$  & $\pm7.5$  & $\pm10.5$  \\ 
    Power cons. (W) & 6.9 & 6.9 & 6.9 & 6.9 & 6.9 \\
\hline \hline
\end{tabular}
\caption{Layout of the HPS Test SVT.  Note the elimination of one layer and 90-degree stereo, and the use of only small, single-sensor-sized double-sided modules both above and below he dead zone. }
}
\label{tab:trk} 
\end{table}
\end{center}
%=======================
\begin{center}
\begin{figure}[htp]
    \includegraphics[width=7cm]{figures/HPS_nochamber}
\caption{A rendering of the Test SVT showing the modules on their support plates held by the hinged C-support on the left and the motion levers on the right. The sensors are shown in red and the hybrids in green. The beam enters from the right through a vacuum box with flanges for services. }
\label{fig:tracker_model}
\vspace*{-5mm}
\end{figure}
\end{center}
%=======================
The SVT is built in two separate halves that are mirror reflections of one another about the plane of the 
nominal electron beam.  Each half consists of five, double-sided modules mounted on a support plate that 
provides services to the modules and allows them to be moved as a group relative to the dead zone. The 
two halves of the tracker are connected to hinges mounted on a C-shaped support just beyond layer 5 that 
defines the nominal spacing between the upper and lower halves of the tracker.  A shaft attached to each 
support plate in front of layer 1 extends upstream and connects to a linear shift that transfers motion into the 
vacuum box through bellows to open and close the tracker around the dead zone. The C-support is 
mounted to an aluminum baseplate that defines the position of the SVT with respect to the vacuum 
chamber. Figure~\ref{fig:tracker_halves} shows a photograph of both completed detector halves prior to 
final assembly. 
%=======================
\begin{figure}[htp]
\begin{center}
    \includegraphics[width=7cm]{figures/2012-101-PHOTON-DETECTOR-001}
\caption{\small{Both halves of the HPS Test SVT after final assembly at SLAC.  The cooling manifolds and 
integrated cable runs are clearly seen.} }
\label{fig:tracker_halves}
\end{center}
\vspace*{-5mm}
\end{figure}
%=======================

@@

\subsubsection{Silicon Sensors}
The sensors for the SVT are selected according to a number of requirements.  First, the sensors must be 
radiation tolerant to approximately $1.5\times10^{14}$~\fluenceunit for a six month run.  This 
corresponds to about 4~MHz/mm$^2$, so the sensor and readout technology must be capable of handling 
very high local occupancies. Since sensitivity is limited by multiple scattering, a material budget of less 
than 1\% $X_0$/layer is imperative and far less is desirable. To achieve the required vertex resolution  
single-hit resolutions better than XXXX~$\mu$m are required. Finally, the sensor technology must 
be mature and readily available at low cost.  Silicon microstrips are the best match to these requirements, and 
a supply of appropriate microstrip sensors purchased from the Hamamatsu Photonics Corporation for the 
Run 2b upgrade of D\O~\cite{Denisov:2001aa} are readily available. 
These are $p+$-on-$n$, single sided, AC coupled, 
polysilicon-biased sensors fabricated on $<100>$ silicon and have 30(60) micron sense(readout) pitch 
over their 4 cm 
$\times$ 10 cm surface. Although a maximum bias of 350 V is specified, the vast majority are operable to 
1000V and therefore remain fully depleted to a dose of approximately 
$1.5\times10^{14}$~\fluenceunit.

\subsubsection{Front-end Electronics}
Because the regions of high occupancy are small spots in two dimensions, only a short length of any one 
strip sees significant occupancy, and the strips in that region act as long pixels. However, the rates are still 
very high and lowering the peak occupancy in the sensors to approximately 1\% for tracking requires a 
trigger window and tagging of hit times to roughly 8~ns, as shown in Fig.~\ref{fig:rates}.  The FADC 
readout for the ECal and muon system are capable of this, but achieving achieving $2\sigma$ efficiency for 
silicon hits then requires 2~ns time resolution for the hits in the SVT. This performance can be achieved with 
the APV25 readout ASIC developed for the CMS experiment at CERN~\cite{French:2001xb}.  When 
operated in ``multi-peak mode", the APV25 captures successive samples of the output of the shaper in sets 
of three.  By fitting the known $CR$-$RC$ shaping curve to these samples, the initial time of the hit can be 
determined to a precision of 2~ns for S/N$>25$, an achievable figure with our sensors if read out 
individually~\cite{Friedl:2009zz}.  For the SVT, six-sample readout and the shortest possible shaping time 
(35~ns) will be used to best distinguish hits that overlap in time.
The APV25 chips are hosted on simple FR4 hybrids since these hybrids and their cooling system are 
outside the tracking volume.  Along with the sensors, these are assembled into half-modules, one for each 
tracking view, with a short cable pigtail to provide power and configuration to the modules and signal 
readout.

@@

The sensor modules for the SVT consist of a pair of identical half-modules, sandwiched back-to-back around 
an aluminum cooling block at one end and a similar PEEK spacer block at the other. 
Figure~\ref{fig:tracker_module} shows a single module after assembly.
%=======================
\begin{figure}[htp]
	\begin{center}
   	 \includegraphics[width=7cm]{figures/IMG_5200}
	\caption{\small{A prototype module assembly (foreground) with the 50~mrad (left) and 100~mrad (right) module assembly fixtures in the background.  A pair of cooling blocks and a spacer block can be seen on the fixtures.} }
	\label{fig:tracker_module}
	\end{center}
\vspace*{-5mm}
\end{figure}
%=======================
The cooling block provides the primary mechanical support for the module as well as cooling via copper 
tubes pressed into grooves in the plates. The spacer block defines the spacing between the sensors at the 
far end of the module, stiffens the module structure, and improves the stability of the sensor alignments.
